Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses

### 2021-04-04 manuscript/manuscript.Rmd
Manuscript knits to all outputs now

---
title: "Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "April 3, 2021"
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/canary-ngos
abstract: >-
  Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. This new wave of attempts to repress civil society organization-many of which are sanctioned by law—represent a bureaucratic form of repression that could indicate more severe human rights abuses and democratic backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do crackdowns on NGOs predict broader human rights repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of political terror, physical violence, and violations of private civil liberties. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society. 
keywords: "civil society restrictions, human rights, political terror, civil liberties"
reference-section-title: References
bibliography: bibliography.bib
title-page: true
editor_options: 
  chunk_output_type: console
---

@@

tribble(
  ~Measure, ~Category, ~Description, ~Source,
  "Political Terror Scale (PTS)", NA, "Ordered category ranging from 1 (least terror) to 5 (most terror)", "(ref:pts)",
  "Physical violence index", NA, "0–1; higher values indicate less violence", "(ref:vdem); v2x_clphy",
  "Private civil liberties index", NA, "0–1; higher values indicate greater respect for civil liberties", "(ref:vdem); v2x_clpriv",
  "Total legal barriers", "De jure laws", "Count of all anti-NGO legal barriers", "(ref:chaudhry)",
  "Barriers to advocacy", "De jure laws", "Count of laws restricting NGO advocacy", "(ref:chaudhry)",
  "Barriers to entry", "De jure laws", "Count of laws restricting NGO entry, including registration", "(ref:chaudhry)",
  "Barriers to funding", "De jure laws", "Count of laws restricting NGO funding sources", "(ref:chaudhry)",
  "CSO repression", "De facto repression", "≈ −3–3; higher values indicate less repression", "(ref:vdem); v2csreprss"
) %>%
  kbl(caption = "Outcomes and predictors",
      booktabs = TRUE, linesep = "") %>% 
  kable_styling(font_size = 8.5) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(1, width = "0.14\\\\textwidth") %>% 
  column_spec(2, width = "0.14\\\\textwidth") %>% 
  column_spec(3, width = "0.5\\\\textwidth") %>% 
  column_spec(4, width = "0.2\\\\textwidth") %>% 
  pack_rows(index = c("Outcomes" = 3, "Predictors" = 5))
```

@@

pts_pred_table %>% 
  kbl(align = "lcccccc",
      booktabs = TRUE, linesep = "",
      caption = "Counts of correctly and incorrectly predicted cases across different PTS models") %>% 
  kable_styling(font_size = 8.5) %>%
  kable_styling() %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(1, width = "0.05\\\\textwidth") %>% 
  column_spec(2:7, width = "0.14\\\\textwidth")
```

@@

tbl_rmse %>% 
  kbl(align = "lcc",
      booktabs = TRUE, linesep = "",
      caption = "Root mean squared error across models predicting physical violence and private civil liberties") %>% 
  pack_rows(index = c("Physical violence index" = 6, "Private civil liberties index" = 6)) %>% 
  kable_styling()
```

@@

```{r tbl-improved-pts}
tbl_improved_pts <- tibble::tribble(
               ~Country, ~Year,   ~Actual, ~`Baseline prediction`, ~`Total NGO barriers included`, ~`Civil society repression included`,
            "Guatemala", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
               "Kosovo", 2012L, "Level 1", "Level 2",                    "Level 1",                          "Level 1",
         "Sierra Leone", 2011L, "Level 3", "Level 2",                    "Level 3",                                "—",
  "Congo - Brazzaville", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Somalia", 2013L, "Level 5", "Level 4",                    "Level 5",                                "—",
              "Bahrain", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Bahrain", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
                "Libya", 2011L, "Level 4", "Level 5",                          "—",                          "Level 4"
  )

tbl_improved_pts %>% 
  kbl(align = "llcccc",
      booktabs = TRUE, linesep = "\\addlinespace",
      caption = "Country-year cases where including civil society restrictions improved PTS predictions") %>% 
  kable_styling(latex_options = c("hold_position")) %>% 
  row_spec(0, bold = TRUE) %>% 
  column_spec(1, width = "0.1\\\\textwidth") %>% 
  column_spec(2, width = "0.05\\\\textwidth") %>% 
  column_spec(3:6, width = "0.16\\\\textwidth")
```

@@

modelsummary(models_paper_pts,
             output = "kableExtra",
             statistic = "[{conf.low}, {conf.high}]",
             coef_map = coef_list,
             gof_omit = "ELPD",
             escape = FALSE,
             caption = "Full results from ordered logistic regression models predicting political terror",
             notes = list("Posterior means; 95% credible intervals in brackets")) %>% 
  kable_styling(font_size = 6) %>% 
  column_spec(1, width = "0.10\\\\textwidth") %>% 
  column_spec(2:6, width = "0.14\\\\textwidth")
```

@@

modelsummary(models_paper_clphy,
             output = "kableExtra",
             statistic = "[{conf.low}, {conf.high}]",
             coef_map = coef_list,
             gof_omit = "ELPD",
             escape = FALSE,
             caption = "Full results from Gaussian regression models predicting physical violence",
             notes = list("Posterior means; 95% credible intervals in brackets")) %>% 
  kable_styling(font_size = 7.5) %>% 
  column_spec(1, width = "0.10\\\\textwidth") %>% 
  column_spec(2:6, width = "0.14\\\\textwidth")
```

@@

modelsummary(models_paper_clpriv,
             output = "kableExtra",
             statistic = "[{conf.low}, {conf.high}]",
             coef_map = coef_list,
             gof_omit = "ELPD",
             escape = FALSE,
             caption = "Full results from Gaussian regression models predicting private civil liberties",
             notes = list("Posterior means; 95% credible intervals in brackets")) %>% 
  kable_styling(font_size = 7.5) %>% 
  column_spec(1, width = "0.10\\\\textwidth") %>% 
  column_spec(2:6, width = "0.14\\\\textwidth")
```

>>>>

---
title: "Canary in the Coal Mine? Civil Society Repression as a Predictor of Human Rights Abuses"
short-title: "Canary in the Coal Mine"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  email: schaudhry@lclark.edu
  url: http://www.suparnachaudhry.com/
- name: Andrew Heiss
  affiliation: Georgia State University
  email: aheiss@gsu.edu
  url: https://www.andrewheiss.com/
date: "April 3, 2021"
published: This is an early draft of a working paper.
git-repo: https://github.com/andrewheiss/canary-ngos
abstract: >-
  Many countries across the globe are increasingly cracking down on civil society groups through legal restrictions. This new wave of attempts to repress civil society organizations—many of which are sanctioned by law—represent a bureaucratic form of repression that could indicate more severe human rights abuses and democratic backsliding. This is especially the case for electoral democracies, which unlike autocracies do not aggressively and openly attack civic space. In this paper, we ask: do crackdowns on NGOs predict broader human rights repression? We argue that this is the case because anti-NGO laws are among the most subtle means of repression and attract the least least domestic and international condemnation, as many audiences mistake this repression as regulation and are not personally affected by it. These laws also make it easier for states to prevent and deter future large-scale popular mobilization. Using original data on civil society restrictions over the last three decades, we test whether NGO crackdown is a predictor of political terror, physical violence, and violations of private civil liberties. This project has important implications: if crackdown on NGOs is indeed an early warning, it can potentially prevent further deteriorations in space for civil society. 
keywords: "civil society restrictions, human rights, political terror, civil liberties"
reference-section-title: References
bibliography: bibliography.bib
title-page: true
endnotes: false
endfloat: false
mainfont: Cochineal
sansfont: Inter
editor_options: 
  chunk_output_type: console
---

@@

tbl_var_summary_caption <- "Outcomes and predictors"

tbl_var_summary <- tribble(
  ~Measure, ~Category, ~Description, ~Source,
  "Political Terror Scale (PTS)", NA, "Ordered category ranging from 1 (least terror) to 5 (most terror)", "(ref:pts)",
  "Physical violence index", NA, "0–1; higher values indicate less violence", "(ref:vdem); v2x_clphy",
  "Private civil liberties index", NA, "0–1; higher values indicate greater respect for civil liberties", "(ref:vdem); v2x_clpriv",
  "Total legal barriers", "De jure laws", "Count of all anti-NGO legal barriers", "(ref:chaudhry)",
  "Barriers to advocacy", "De jure laws", "Count of laws restricting NGO advocacy", "(ref:chaudhry)",
  "Barriers to entry", "De jure laws", "Count of laws restricting NGO entry, including registration", "(ref:chaudhry)",
  "Barriers to funding", "De jure laws", "Count of laws restricting NGO funding sources", "(ref:chaudhry)",
  "CSO repression", "De facto repression", "≈ −3–3; higher values indicate less repression", "(ref:vdem); v2csreprss"
)

if (!is_docx) {
  tbl_var_summary %>%
    kbl(caption = tbl_var_summary_caption,
        booktabs = TRUE, linesep = "") %>% 
    pack_rows(index = c("Outcomes" = 3, "Predictors" = 5)) %>% 
    kable_styling(font_size = ifelse(is_latex, 8.5, NA)) %>% 
    row_spec(0, bold = TRUE) %>% 
    column_spec(1, width = "0.14\\\\textwidth") %>% 
    column_spec(2, width = "0.14\\\\textwidth") %>% 
    column_spec(3, width = "0.5\\\\textwidth") %>% 
    column_spec(4, width = "0.2\\\\textwidth")
} else {
  tbl_var_summary %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    insert_row("Outcomes", "", "", "", after = 1) %>% 
    insert_row("Predictors", "", "", "", after = 5) %>% 
    set_bold(1, everywhere) %>%
    set_bold(c(2, 6), everywhere) %>%
    set_bottom_border(c(2, 6), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_var_summary_caption) %>% 
    huxtable::as_flextable()
}
```

@@

pts_pred_table_caption <- "Counts of correctly and incorrectly predicted cases across different PTS models"

if (!is_docx) {
  pts_pred_table %>% 
    kbl(align = "lcccccc",
        booktabs = TRUE, linesep = "",
        caption = pts_pred_table_caption) %>% 
    kable_styling(font_size = ifelse(is_latex, 8.5, NA)) %>%
    kable_styling() %>% 
    row_spec(0, bold = TRUE) %>% 
    column_spec(1, width = "0.05\\\\textwidth") %>% 
    column_spec(2:7, width = "0.14\\\\textwidth")
} else {
  pts_pred_table %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_bold(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    # set_col_width(1, 0.05) %>%
    # set_col_width(2:7, 0.14) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(pts_pred_table_caption) %>% 
    huxtable::as_flextable()
}
```

@@

tbl_rmse_caption <- "Root mean squared error across models predicting physical violence and private civil liberties"

if (!is_docx) {
  tbl_rmse %>% 
    kbl(align = "lcc",
        booktabs = TRUE, linesep = "",
        caption = tbl_rmse_caption) %>% 
    pack_rows(index = c("Physical violence index" = 6, "Private civil liberties index" = 6)) %>% 
    kable_styling()
} else {
  tbl_rmse %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1, "left") %>% 
    set_align(everywhere, 2:3, "center") %>% 
    insert_row("Physical violence index", "", "", after = 1) %>% 
    insert_row("Private civil liberties index", "", "", after = 8) %>% 
    set_bold(1, everywhere) %>%
    set_bold(c(2, 9), everywhere) %>%
    set_bottom_border(c(2, 9), everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_rmse_caption) %>% 
    huxtable::as_flextable()
}
```

@@

```{r tbl-improved-pts}
tbl_improved_pts <- tribble(
               ~Country, ~Year,   ~Actual, ~`Baseline prediction`, ~`Total NGO barriers included`, ~`Civil society repression included`,
            "Guatemala", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
               "Kosovo", 2012L, "Level 1", "Level 2",                    "Level 1",                          "Level 1",
         "Sierra Leone", 2011L, "Level 3", "Level 2",                    "Level 3",                                "—",
  "Congo - Brazzaville", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Somalia", 2013L, "Level 5", "Level 4",                    "Level 5",                                "—",
              "Bahrain", 2011L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
              "Bahrain", 2013L, "Level 3", "Level 2",                    "Level 3",                          "Level 3",
                "Libya", 2011L, "Level 4", "Level 5",                          "—",                          "Level 4"
  )

tbl_improved_pts_caption <- "Country-year cases where including civil society restrictions improved PTS predictions"

if (!is_docx) {
  tbl_improved_pts %>% 
    kbl(align = "llcccc",
        booktabs = TRUE, linesep = "\\addlinespace",
        caption = tbl_improved_pts_caption) %>% 
    kable_styling(latex_options = c("hold_position")) %>% 
    row_spec(0, bold = TRUE) %>% 
    column_spec(1, width = "0.1\\\\textwidth") %>% 
    column_spec(2, width = "0.05\\\\textwidth") %>% 
    column_spec(3:6, width = "0.16\\\\textwidth")
} else {
  tbl_improved_pts %>% 
    hux() %>% 
    set_font_size(10) %>%
    set_all_padding(3) %>% 
    set_align(everywhere, 1:2, "left") %>% 
    set_align(everywhere, 3:6, "center") %>% 
    set_bold(1, everywhere) %>%
    set_bottom_border(1, everywhere) %>%
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_improved_pts_caption) %>% 
    huxtable::as_flextable()
}
```

@@

tbl_reg_pts_caption <- "Full results from ordered logistic regression models predicting political terror"

tbl_reg_pts <- modelsummary(models_paper_pts,
                            output = table_format,
                            statistic = "[{conf.low}, {conf.high}]",
                            coef_map = coef_list,
                            gof_omit = "ELPD",
                            escape = FALSE,
                            caption = tbl_reg_pts_caption,
                            notes = list("Posterior means; 95% credible intervals in brackets"))

if (!is_docx) {
  tbl_reg_pts %>% 
    kable_styling(font_size = ifelse(is_latex, 7, NA)) %>% 
    column_spec(1, width = "0.10\\\\textwidth") %>% 
    column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_pts %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    # set_col_width(1, 0.1) %>% 
    # set_col_width(2:6, 0.14) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_pts_caption) %>% 
    huxtable::as_flextable()
}
```

@@

tbl_reg_clphy_caption <- "Full results from Gaussian regression models predicting physical violence"

tbl_reg_clphy <- modelsummary(models_paper_clphy,
                              output = table_format,
                              statistic = "[{conf.low}, {conf.high}]",
                              coef_map = coef_list,
                              gof_omit = "ELPD",
                              escape = FALSE,
                              caption = tbl_reg_clphy_caption,
                              notes = list("Posterior means; 95% credible intervals in brackets"))

if (!is_docx) {
  tbl_reg_clphy %>% 
    kable_styling(font_size = ifelse(is_latex, 7.5, NA)) %>% 
    column_spec(1, width = "0.10\\\\textwidth") %>% 
    column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_clphy %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_clphy_caption) %>% 
    huxtable::as_flextable()
}
```

@@

tbl_reg_clpriv_caption <- "Full results from Gaussian regression models predicting private civil liberties"

tbl_reg_clpriv <- modelsummary(models_paper_clpriv,
                               output = table_format,
                               statistic = "[{conf.low}, {conf.high}]",
                               coef_map = coef_list,
                               gof_omit = "ELPD",
                               escape = FALSE,
                               caption = tbl_reg_clpriv_caption,
                               notes = list("Posterior means; 95% credible intervals in brackets"))

if (!is_docx) {
  tbl_reg_clpriv %>% 
    kable_styling(font_size = ifelse(is_latex, 7.5, NA)) %>% 
    column_spec(1, width = "0.10\\\\textwidth") %>% 
    column_spec(2:6, width = "0.14\\\\textwidth")
} else {
  tbl_reg_clpriv %>% 
    set_font_size(10) %>% 
    set_all_padding(1) %>% 
    set_align(everywhere, 2:6, "center") %>% 
    set_bold(1, everywhere) %>% 
    set_width(1) %>% 
    set_wrap(TRUE) %>% 
    huxtable::set_caption(tbl_reg_clpriv_caption) %>% 
    huxtable::as_flextable()
}
```
